---
title: "Cyclistic - A Case Study"
author: "Prasanna Sriram"
date: "2024-09-03"
output: 
  github_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

A case study to answer business questions of a fictional bike sharing company - Cyclistic. The goal of the study is to answer questions raised by Marketing team.

The main question to analyze was how the annual members and casual riders used the Cyclistic bikes differently.


### **Data Collection and Preparation**

We will be using **Cyclistic's historical data**. This is publicly available data. We verified the integrity and privacy of the data collected. The key tasks that were performed during this step was: 

* Download the data
* Identify how it's organized
* Sort and filter the data
* Determine the credibility of the data

### **Data Processing**

In this step, we processed the data using **R Studio**. The following steps were performed

* Checked data for errors. We noted that the data from different files were stored under different column names.
* We noticed that some of the data were stored as different data types.
* Removed unwanted columns in the datasets.
* Combined all the datasets to make a single dataset.
* Transformed data in certain columns to remove repeated values.
* Created separate columns for year, month and days to analyze the data more effectively.
* Removed bad data which had negative ride length values.

```{r all_trips_summary, echo=FALSE}
all_trips_summary <- read.csv("all_trips_summary.csv")
all_trips_summary
```

### **Data Analysis**

The key tasks for this stage were

* Aggregation of data so that it's useful and accessible.
* Organizing and formatting of the data.
* Performing the necessary calculations.
* Identifying trends and relationships.

We primarily analyzed two factors

#### **Average Ride length of Casual v/s Members by day**

```{r all_trips_avg_ride_time, echo=FALSE}
all_trips_avg_ride_time <- read.csv("all_trips_avg_ride_time.csv")
all_trips_avg_ride_time
```

#### **Average Ride time of Casual v/s Members by day**


```{r rider_data_by_type_week, echo=FALSE}
rider_data_by_type_week <- read.csv("rider_data_by_type_week.csv")
rider_data_by_type_week
```


### **Data Visualizations**

You can also embed plots, for example:

```{r number_of_rides, echo=FALSE}
plot1 <- read.csv('plot1_data.csv')
plot1 %>% ggplot(aes(x = weekday, y = number_of_rides, fill = member_casual)) +
  geom_col(position = "dodge")
```

```{r average_duration, echo=FALSE}
plot2 <- read.csv('plot2_data.csv')
plot2 %>% ggplot(aes(x = weekday, y = number_of_rides, fill = member_casual)) +
  geom_col(position = "dodge")
```
